\chapter{Introduction}

Quality assurance (QA) is an easily overlooked process in many companies.  A good quality assurance (QA) system increases customer satisfaction, credibility of the company and optimizes development flows. Solid QA is a fundamental and necessary step in every respectable software engineering project.\\\\
Failure to maintain a proper quality assurance program can lead to catastrophic disasters. On January 28, 1986, The Space Shuttle Challenger exploded, destroying the vehicle and killing several crew members. This accident was caused by a hardware failure of a solid rocket booster. Several severe flaws in quality assurance and human decision making contributed to this event.\\\\
Quality assurance is not limited to software. Toyota Motor Corp recalled 3.8 million vehicles in 2009 due to problems with a removable floor mat that could cause accelerators to get stuck. This issue can cause a crash, serious injuries or even death.\\\\
Software bugs can even lead to lost customer money, as the recent DAO vulnerabilities has demonstrated [https://blog.ethereum.org/2016/06/17/critical-update-re-dao-vulnerability/]. On June 17, 2016, Etherium, a blockchain-based decentralized platform for applications, reported a vulnerability in their code, allowing an attacker to steal small sums of cryptocurrency. Proper testing, code review processes and quality assurance infrastructures might have prevented these events.\\\\
The entire software development cycle, from design to deployment, is involved with quality assurance. This includes software design, code reviews, testing, version control systems, release management and customer interaction.\\\\
Tribler is the result of 10 ongoing years of scientific research. The software allows users to discover and download interesting content such as music and videos. Tribler is built upon the \emph{libtorrent} library. In 2014, support for anonymous download using a Tor-like protocol has been added.\\\\
The content of this thesis is structured as follows. The first chapter (linkje) explains ............

% Improvement of Key Problems of Software Testing in Quality Assurance: https://arxiv.org/pdf/1202.2506.pdf